---
title: "Homework 1 PH 290"
author: "Nolan Gunter"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---
```{r}
library(assertthat)
library(ggplot2)
library(testthat)
library(tidyverse)
library(ggpubr)
```

# 1. Write down two differences between the type and the class of an R object

An R object's `type` denotes its internal storage mode, while we can view `class` as more of an external form of an object that demonstrates how functions should interpret it. For example, consider the following:

```{r}
x <- 4.56
y <- 4L
typeof(x); typeof(y)
class(x); class(y)
```

There is a nuance between the two components of objects x and y. Type denotes a more specific case of x, making it specified as a double instead of strictly as a numeric to the R S3 method dispatch. However, type and class are technically the same for our integer y. 

# 2. Create a function `fp` to calculate the $p^{th}$ absolute power of a vector of numbers. 

The following function `fp` calculates the absolute power of a numeric value, or a vector of numerics by default. The inputs include the numeric(s) value(s), the power p which must be greater than or equal to 1. If the user sets `deriv = TRUE`, then the derivative will be calculated instead. 

```{r}
fp <- function(x,p = 2,deriv = FALSE){
  
  #p must be greater than or equal to 1, simultaneously checks it is numeric
  stopifnot(p >= 1)
  
  stopifnot(is.numeric(x))
  
  if(deriv){
    if(p == 1) sign(x)
    else p*(abs(x)^(p-1))*sign(x)
  }
  else{
    abs(x)^p
  }
}

test_that("fp works", {
  a <- c(1,2,3)
  b <- c(-1,2,-3)
  c <- as.numeric(c())
  
  expect_equal(fp(a, p=2), c(1,4,9))
  expect_equal(fp(a, p=4), c(1,16,81))
  expect_equal(fp(b, p=1), c(1,2,3))
  expect_equal(fp(b, p=2, deriv=TRUE), c(-2,4,-6))
  expect_equal(fp(b, p=1, deriv=TRUE), c(-1,1,-1))
  expect_equal(fp(c), numeric(0))
})
```

\newpage
# 3. Subgroup analyses on diabetes data

```{r}
diabetes <- read.csv(file = "~/Desktop/diabetic_data.csv") %>% 
  select(race, gender, A1Cresult, time_in_hospital) %>%
  filter(gender != "Unknown/Invalid")
  
##created nested data based on gender and race
nest_diabetes <- diabetes %>% group_by(gender, race) %>% tidyr::nest()

final_data <- nest_diabetes %>%
              mutate(ct = map(data, function(dat) with(dat,table(A1Cresult, 
                       time_in_hospital, useNA = "ifany"))),
                     tt = map(data, function(dat) t.test(
                       dat %>% filter(A1Cresult %in% c("Norm")) %>% pull(time_in_hospital),
                       dat %>% filter(A1Cresult %in% c(">7", ">8")) %>% pull(time_in_hospital)
                     )),
                     pval = map_dbl(tt, function(t) t$p.value)
                  )

final_data[1:5,c("race","gender","pval")]
```

\newpage
# 4. Design and write a utility function to compare linear regression models side-by-side. 

The following function `compareLM` is designed to facilitate quick comparison of linear models. The user must input the dataset, the outcome/response variable (that will be the same across models), and a list of character vectors representing the variables (with names as they appear in the data) requesting to be used in the data. 

The function relies on a sub-function called `getLM`, which creates a linear model from the data.

```{r}
getLM <- function(outcome, vars, dat){
  lin <- lm(data = dat, 
            formula = paste(outcome, "~", 
                      paste(vars, collapse = " + ")))
}

compareLM <- function(data, outcome, var_list){
  
  #ensure outcome variable defined in dataset
  assert_that(any(outcome == colnames(data)))
  
  #create new data with outcome observed values
  new_dat <- data[,outcome]
  
  assert_that(length(var_list) >= 1)
  num_mods <- length(var_list)
  
  #create list for summary objects
  lm_summaries <- list()
  
  #for loop for linear model, prediction and summary info
  for(i in 1:num_mods){
    lm_tmp <- getLM(outcome, var_list[[i]], data)
    new_dat <- cbind(new_dat,predict(lm_tmp, newdata = gap))
    lm_summaries[[i]] <- summary(lm_tmp)
  }
  
  colnames(new_dat) <- c("Observed", paste0("Predict", 1:num_mods))
  
  #create plots to merge for each individual model
  for(i in 1:num_mods){
    plot <- ggplot(new_dat, aes_string(x = "Observed", y = paste("Predict", i, sep = ""))) +
      geom_point() + geom_smooth(method = "lm") 
    assign(paste0("plot", i), plot)
  }
  
  #merge plots using ggarrange function in ggpubr
  p <- paste(paste0("plot",1:num_mods), collapse = ",")
  ret_plot <- eval(parse(text=paste("ggarrange(", p, ")", sep = "")))
  
  return(list(new_dat, ret_plot, lm_summaries))
}
```

Now we test the code on the gapminder dataset. 

```{r}
gap <- gapminder::gapminder
outcome <- "lifeExp"
var_list <- list(c("year"), 
                 c("year", "gdpPercap"), 
                 c("year", "gdpPercap", "pop"), 
                 c("year", "gdpPercap", "pop", "continent"))

new <- compareLM(data = gap, outcome, var_list)
head(new[[1]])
new[[2]]
new[[3]]
```

Outside sources: 

https://stackoverflow.com/questions/17024685/how-to-use-reference-variables-by-character-string-in-a-formula

https://cran.r-project.org/web/packages/jtools/vignettes/summ.html

https://stackoverflow.com/questions/15272916/how-to-wait-for-a-keypress-in-r

