---
title: "PH290_HW1"
output:
  pdf_document: default
  html_document: default
date: '2022-09-12'
---
Noah Baker
2022-09-13


```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(testthat)
library(purrr)
```



## 1. Write down two differences between the type and the class of an R object

The type is the storage mode of an object while the class is an attribute of an object, in object-oriented programming, that names the classes the object inherits.

Types cannot be changed unless changing the storage of the object while variable classes can be renamed.


## 2. Create a function fp to calculate the pth absolute power of a vector of numbers

```{r}
fp <- function(x, p=2, deriv=FALSE){
  if(deriv == TRUE){
    if(p>1){
      p*abs(x)**(p-1)*sign(x)
    }
    else{
      sign(x)
    }
  }
  else{
    abs(x)**p
  }
}  
```

```{r}
test_that("fp works", {
  x <- 6
  y <- -4
  t.1 <- c(2, 3, 4)
  t.2 <- 4
  t.3 <- 1
  t.4 <- 2
  t.5 <- numeric(length=0)
  expect_equal(fp(t.1), c(4, 9, 16))
  expect_equal(fp(x, t.2), abs(x)**4)
  expect_equal(fp(y, t.3), abs(y))
  expect_equal(fp(x, t.4, deriv = TRUE), 2*x)
  expect_equal(fp(y, t.3, deriv = TRUE), sign(y))
  expect_equal(fp(t.5), numeric(length=0))
})
```


## 3. Perform a subgroup analyses using a diabetes dataset

```{r}
# this reads in the diabetes dataset, selects the given columns, and reassigns
# categorical variables to binary values of abnormal and normal, then cleans 
# the data for all values missing

diab = read.csv("data/dataset_diabetes/diabetic_data.csv") %>%
  select(race, gender, A1Cresult, time_in_hospital) %>%
  mutate(A1Cresult = case_when(
    A1Cresult == ">7" | A1Cresult == ">8" ~ "Abnormal",
    A1Cresult == "Norm" ~ "Normal",
    TRUE ~ "None"
  ))
diab_slim = diab  %>% filter(.$A1Cresult != "None")
```


```{r}
# this function creates a dataframe with rows grouped by gender and race. 
# Subset within each group, a cross tabulation and t-test for normal vs abnormal
# alcohol levels and time spent in the hospital

subgroups <- diab_slim %>%
  group_by(gender, race) %>%
  tidyr::nest() %>%
  mutate(
    table <- map(data, function(dat){
      xtabs(~A1Cresult + time_in_hospital, dat)
    }),
    ttest <- map(data, function(x){
      t.test(time_in_hospital ~ A1Cresult, data = x)
    })
  )
```

```{r}
# outputs of the results of subgroups' cross tabulation and statistical t-test 
# values

for(n in 1:3){
  print(paste0(map(subgroups[[1]], head)[[n]]," ", map(subgroups[[2]], head)[[n]]))
  print(map(subgroups[[4]], head)[[n]])
  print(map(subgroups[[5]], head)[[n]])
}
```


## 4. Design and write a utility function to compare linear regression models side-by-side.

```{r}
# these create the linear regression models to be compared
mod1 <- lm(lifeExp ~ year, gapminder::gapminder)
mod2 <- lm(lifeExp ~ year + gdpPercap, gapminder::gapminder)

mods <- list(mod1, mod2)

# function that takes in a list of linear models and prints their coefficients 
# and produces graphical comparisons
compare_mod <- function(mods) {
  # prints the coefficients of the linear regresion models
  coeff <- map(mods, function(m) {coefficients(m)})
  # produces graphs for each of the mods linear regression models
  graphs <- map(mods, function(m) {
    # plots with a 2x2 grid
    par(mfrow=c(2,2))
    # plots the functions
    plot(m)
    })
  return(c(coeff, graphs))
}

# visualization
compare_mod(mods = mods)
```
























