---
output: rmdformats::html_clean
title: "Biostatistical Computing: Homework 1"
author: Madeline Adee
date: "`r format(Sys.time(), '%d %B, %Y')`"
---


Initial setup and package loading. 

```{r setup, include=TRUE}

knitr::opts_chunk$set(echo = TRUE)
# use pacman package to load packages (easier when switching between computers)
pacman::p_load(testthat, # testing package
               dplyr, # data cleaning, pipes, etc.
               purrr, # functional programming
               tidyr,
               kableExtra, # nice HTML tables for results
               expss, # apply nice labels to data
               medicaldata, # sample dataset
               ggplot2, # plotting
               ggfortify, # allow plotting of lm objects with ggplot
               gridExtra, # to add title to ggfortify plot
               broom, 
               gtsummary
)

```

# Problem 1

Two differences between the type and the class of an R object are: 

- The type of an object refers to the internal storage mode (can be accessed using `typeof()`) and class of an object determines how other functions interpret the object (can be accessed using `class()`). 
- Class is an attribute of the object. R objects are supplemented with attributes, one of which is class. 


# Problem 2

First, I wrote the function as detailed in the assignment description. 

```{r}

fp <- function(x, p, deriv = FALSE){
  
  # check that inputs are correct format
  stopifnot(is.numeric(x))
  stopifnot(is.numeric(p))
  stopifnot(is.logical(deriv))
  
  if(deriv == FALSE){
  val <- abs(x)^p
  }
  
  if(deriv == TRUE){
  val <- p*abs(x)^(p-1)*sign(x)
  }
  
  return(val)
  
}

```

Then I wrote the tests for the function using testthat::expect_equal, and grouedp the tests together using testthat::test_that. 

```{r}

test_that("fp works", {
  
  x <- c(1, 2)
  x_empty <- numeric(length=0)
  expect_equal(length(fp(x, 2)), length(x))
  expect_equal(fp(x, 4), x^4)
  expect_equal(fp(x, 1), abs(x))
  expect_equal(fp(x, 2, deriv = TRUE), 2*x)
  expect_equal(fp(x, 1, deriv = TRUE), sign(x))
  # what is the point of this one?
  expect_equal(fp(x_empty, 1), x_empty)
  
})

```

# Problem 3

First I loaded and cleaned the data in order to complete the rest of the problem.

```{r}

data <- read.csv("./dataset_diabetes/diabetic_data.csv")

data_select <- data %>%
  # choosing only variables I need for this assignment
  select(race, gender, A1Cresult, time_in_hospital) %>%
  # remove if no A1Cresults since we are comparing two types of A1Cresults results
  filter(A1Cresult != "None") %>%
  # convert ? to "unknown" for nicer looking tables etc.
  mutate(race = ifelse(race == "?", "Unknown", race)) %>%
  # create groupings as specified in assignment
  mutate(A1Cresult = ifelse(A1Cresult == ">7" | A1Cresult == ">8", "Abnormal", A1Cresult), 
         # cleaning up the formatting so this looks nicer
         race = ifelse(race == "AfricanAmerican", "African American", race))

```

Then I wrote the functions needed to run the t-test and crosstabs, and be able to present the results nicely in tables.  

```{r}

# function to run t-test to compare time in hospital for normal vs. abnormal
f_ttest <- function(data) {
  norm <- data %>% filter(A1Cresult == "Norm")
  abnorm <- data %>% filter(A1Cresult == "Abnormal")
  t.test(x = norm$time_in_hospital, y = abnorm$time_in_hospital)
}

# function to run crosstabs and write HTML for table
f_crosstabs <- function(data, label){
 tbl_cross(data, row=time_in_hospital, col=A1Cresult, percent="column", 
           label = list(time_in_hospital ~ "Time in hospital", A1Cresult ~ "A1C Result")) %>%
    modify_header(label = label)
}

# function to save t-test results nicely to make it easier to add table later
f_results <- function(data){

  mean_norm <- data[['estimate']][[1]]
  mean_abnorm <- data[['estimate']][[2]]
  stat <- data[['statistic']][[1]]
  pval <- data[['p.value']] 

  temp_df <- c(
             mean_norm = mean_norm, 
             mean_abnorm = mean_abnorm, 
             stat = stat, 
             pval = pval)
  
  return(temp_df)
}

```

Nest and then use map to generate the t-tests and cross-tabs. 

```{r}

data_grouped <- data_select %>%
  group_by(race, gender) %>%
  tidyr::nest() 

data_summary <- data_grouped %>%
  mutate(t_res = map(data, f_ttest), 
         obs = map(data, function(x) nrow(x)),
         summary_stats = map(t_res, f_results), 
         crosstabs = map(data, ~f_crosstabs(.x, label = paste(race, gender)))) %>%
  tidyr::unnest_wider(summary_stats)

```

Arrange output data for clarity and print out results of crosstabs and t-tests.

```{r, results = "asis"}

# change to alphabetical arrangement of groups by gender then race for clarity
data_summary <- data_summary %>%
  arrange(by = gender, race) 

# printing first two cross-tabs tables only as example (for brevity)
print(data_summary$crosstabs[1:2])

# printing out the t-test results for all groups
data_summary %>%
select(-data, -t_res, -crosstabs) %>%
  kbl(digits = 3, 
      col.names = c("Race", "Gender", "N Observations", 
                    "Normal A1C", "Abnormal A1C", "Test Statistic", "P-value")) %>%
  add_header_above(c(" " = 3, "Mean Days in Hospital" = 2, "T-test Results" = 2)) %>%
  kable_styling()

```

# Problem 4

## Functions

First I wrote a function to plot the diagnostics of the linear models generated. 

```{r}

# function to look at linear model diagnostics plot ============================

f_plot <- function(model, title){
  
p <- autoplot(model, label.size = 3, 
         which = 1:4, ncol = 2,
         colour = '#C6C6C6', 
         smooth.colour = '#00A19C', 
         title = title, 
         ) + 
  theme_bw() 

plot <- gridExtra::grid.arrange(grobs = p@plots, top = title)

return(plot)

}


```

Then I wrote a function to compare linear models. The function allows you to either:

- **compare linear models for many different dependent variables** with the same independent variable specification per model by setting
- **compare multiple linear models different specifications of independent variables** and the same dependent variable per model by setting. For this option you can also specify if you want to look at:
    - all possible combinations of listed independent variables by setting
    - each independent variable on it's own
    
The original model specifications and detailed summary are preserved in the dataframe that is the result of running the function. A short summary of the models can be shown in a table using the dataframe, and diagnostic plots are also accessible in the data frame to be printed. 


```{r}

# Function to compare models ===================================================

f_compare <- function(data, y, x, diff_outcomes = FALSE, type = "comb"){
  
# data: should be a data frame or tibble
  
# y: dependent variable(s) as character string. Should be length of 1 if comparing
# different independent variable specifications, length > 1 if comparing different 
# independent variable specifications
  
# x: independent variable(s) as character string. 
  
# diff_outcomes: TRUE if comparing same specifications for different dependent 
# (outcome) variables, FALSE if comparing different independent variable 
# specifications for same outcome. Defaults to FALSE.
  
# type: only needs to be specified when diff_outcomes = FALSE. This indicates 
# whether you want to compared models with one independent variable per outcome 
# (type = "single") from the input x, or if you want to look at all possible 
# combinations of specifications without interaction from input x. 
  
  
stopifnot(is.character(y))
stopifnot(is.character(x))
stopifnot(is.logical(diff_outcomes))

# create empty model specifications list
model_specs <- c()

# if comparing models for different outcomes ----------------------------------

if(diff_outcomes == TRUE) {
  
  stopifnot(length(y)>1)
  
  # collapse x var for proper formatting
  if (length(x) > 1){
  x <- paste(x, collapse = "+")
  }
  
  # create a list of the model specifications based on input x and y list
  for (out in y) {
  model_specs <- c(model_specs, paste(out, "~", x))
  }
  
}

# if comparing models for same outcome -----------------------------------------
  
if (diff_outcomes == FALSE){
  
  stopifnot(length(y)==1)
  stopifnot(length(x)>1)
  
  if (type == "comb"){
  
  # make list of all possible combinations of x var
  x_list <- do.call("c", lapply(seq_along(x), function(i) combn(x, i, FUN = list)))
  x_list  <- sapply(x_list[!sapply(x_list, purrr::is_empty)], paste, collapse = '+')
  
  }

  if (type == "single") {
  
  x_list <- x
    
  }
  
  # create a list of the model specifications based on input x and y list
  for (item in x_list) {
  model_specs <- c(model_specs, paste(y, "~", item))
  }
  
}

# run the models ---------------------------------------------------------------   

output <-
  tibble(model_specs) %>% 
  # run lm for each specified formula
  mutate(model = map(model_specs, ~lm(formula = .x, data = data)),
         glance = map(model, glance), 
         summary = map(model, summary),
         plot = map(model, ~f_plot(model = .x, title = model_specs))) %>%
  unnest(glance)

return(output)

}

```



## Example scenario 1 
Compare model results for 2 different outcomes. 

Both examples for the function use the `polyps` datset from the `medicaldata` r package. This contains results from a RCT of Sulindac for Polyp Prevention in Familial Adenomatous Poly-posis. The variable **number3m** indicates the number of colonic polyps at 3 months, and **number12m** indicates the number of colonic polyps at 12 months. This is a very small dataset so used for example purposes only. 

```{r}

y <- c("number3m", "number12m")
x <- c("baseline", "age")

model_summary <- f_compare(data = medicaldata::polyps, y = y, x = x, diff_outcomes = TRUE)

model_summary %>%
select(model_specs, r.squared, adj.r.squared, sigma, statistic, p.value) %>%
  kbl(digits = 6, 
      col.names = c("Model", "R Squared", "Adj. R Squared", 
                    "Sigma", "F-Statistic", "P-value")) %>%
  kable_styling()


```


## Example scenario 2
Compare model results for 1 outcome with different model specifications for predictors.

```{r}

y <- c("number3m")
x <- c("baseline", "age", "treatment")
model_summary <- f_compare(data = medicaldata::polyps, y = y, x = x, 
                           diff_outcomes = FALSE, type = "comb")


model_summary %>%
select(model_specs, r.squared, adj.r.squared, sigma, statistic, p.value) %>%
  kbl(digits = 6, 
      col.names = c("Model", "R Squared", "Adj. R Squared", 
                    "Sigma", "Statistic", "P-value")) %>%
  kable_styling()


```

